---
title: "Reproducible Research: Peer Assessment 1"
author: "jalagon"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

First, download the data from internet.

```{r Download and extract data, cache=TRUE}
URL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url = URL, destfile = "activity.zip")
unzip("activity.zip")
```

Now read data in R, change date variable to date format and check everything looks ok.

```{r Read data}
# Read data. I assume you have downloaded zip data and extracted *activity.csv*  to working directory.
activity <- read.csv("activity.csv")
# Change date variable to date format
activity$date<-as.Date.factor(activity$date)
# Explore data frame
str(activity)
head(activity)
tail(activity)

```


## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r Total steps per day}
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
#Calculate total number of steps per day
StepsPerDay <- summarise(group_by(activity, date), sum(steps))
StepsPerDay

```

2. Make a histogram of the total number of steps taken each day
```{r Histogram}
hist(StepsPerDay$`sum(steps)`,
     main = "Total number of steps per day",
     xlab = "Steps per day")
```


3. Calculate and report the mean and median of the total number of steps taken per day
```{r Mean and median steps per day}
#Calculate mean and median across days, removing NA
mean(StepsPerDay$`sum(steps)`, na.rm = TRUE)
median(StepsPerDay$`sum(steps)`, na.rm = TRUE)

```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
